Sizing the effects of temperature on fish

### 2017-09-11 draft.Rnw
arg
final psoter etc

Increasing temperatures under climate change are thought to affect individual physiology of ectotherms through increases in metabolic demands, leading to decreases in body-size with concomitant effects on species ecology. Although intuitively appealing, the magnitude, as well as when and where such changes are expected in an ecosystem context remains unclear. We develop a general trait-based model of carbon and oxygen budgets to determine the relative influence of thermal sensitivity, oxygen limitation, and environmental conditions on ectotherm body size.  Our model provides a physiological basis for changes in species size and observed ecological consequences of temperature, and ties together various strands of thought about the ecological consequences of increased temperatures.

@@

Temperature, through its effects on individual physiology, is a dominant driver of species ecology and bio-geography (\cite{deutsch_climate_2015,pinsky_marine_2013,brown_toward_2004}). As a consequence, current and predicted temperature increases under climate change will act as a strong agent of change in many ecosystems (\cite{deutsch_climate_2015, stuart-smith_thermal_2015}). However, the nature of these changes can be difficult to predict as temperature effects scale from individuals to species and ecosystems. Through this cascade of scales, incorrect or approxiamte model assumptions at the individual scale can have disproportionate effects on ecosystem level outcomes (\cite{brander_overconfidence_2013, lefevre_models_2017}). In marine fish, recent predictions of decreasing organism size, and resulting decreases in the size of fisheries catches (\cite{cheung_shrinking_2013}), have been criticized as overly simplistic and not in line with physiological constraints (\cite{brander_overconfidence_2013, lefevre_models_2017}). While it is not disputed that organism size can change as a result of changing temperatures, the question remains when, where and why this is so.

Empirically, most ectotherms are found to grow larger in colder temperatures - a phenomenon called the temperature-size rule (\cite{atkinson_cells_2006}). Models of ectotherm growth can reproduce this pattern (under assumptions of optimal energy allocation between growth and reproduction) if increases in metabolic costs (in terms of energy) with temperature outpace increases in energy acquisition (\cite{kozlowski_can_2004}). However, while increasing metabolic costs are real, energy acquisition is an abstract, phenomenological description that lumps ecological aspects of foraging activity and prey availability. At a more mechanistic level of energy acquisition, consistent responses of foraging and energy aquisition with temperature have remained elusive (\cite{englund_temperature_2011,rall_universal_2012}), with variable responses across habitats and taxa. This is perhaps not surprising given variable foraging ecologies, but it remains unclear how temperature dependant feeding ecology impacts growth and adult size.

Physiologically, a long held view has been that temperature is a controlling factor while oxygen supply sets the physiological limits (\cite{fry_effects_1947,claireaux_linking_2007,lefevre_are_2016}). How exactly temperature influences physiological rates and limits, however, has been a matter of debate, not least because of the variable responses observed among different species. In most species, the standard metabolism (i.e., the metabolic cost of maintenance and routine activity such as ventilation) increases near exponentially with temperature. A prevalent view is that the maximum metabolic rate (i.e., the metabolic rate at maximum sustained exercise) has a  dome-shaped response to temperature, whereby it can be increased (passively and actively) up to a point, but plateaus or decreases thereafter (\cite{fry_effects_1947,claireaux_linking_2007,lefevre_are_2016,portner_physiology_2008}). This leads to the view of a modal curve for metabolic scope (the available oxygen/energy for activity), and suggests that towards the upper end of this curve, organisms will run out of oxygen.

This view was encapsulated in the Theory of Oxygen and Capacity Limitation of Temperature (\cite{portner_oxygen-and_2010}), which suggests that the decrease in metabolic scope towards extreme temperatures limits species ability to sustain core functions such as foraging and growth. In some species, however, maximum metabolic rate may be steadily increasing (\cite{lefevre_are_2016, verberk_does_2016}), suggesting that oxygen may not be the limiting factor. Consequently, the difference in observed physiological responses may at least partially determine temperature effects on body size for individual species. In addition to physiological differences, the assumed coefficients for temperature sensitivity of metabolism and size scaling exponents are different between species, and can disproportionately affect inferences about future organism size \cite{lefevre_models_2017, kozlowski_can_2004}.

@@

\section{Setting up the model}

@@

Our model assumes that physiology is described by two key budgets: the energy and oxygen budgets. In setting up the model we make a number of key assumptions, which are similar to those made in Holt \& Jorgensen. At a high level, we assume that animals will adapt activity levels to optimise available energy for growth and reproduction. Growth is limited either by food capture, food processing capacity, or by available oxygen. We further assume that temperature acts directly on rates that are determined by enzymatic activity: digestive activity (via maximum consumption) and metabolic costs. The scaling with temperature (determined by the activation energy $E_a$) is assumed the same for these processes. Consequently, temperature only acts on ecological rates (food acquisition) via optimisation given temperature driven changes in enzymatic rates.

@@

The oxygen budget $P_{0_2}$, determines physiological limits to aerobic acitivity. Demand ($D_{O_2}$) is the sum of oxygen used for SDA (i.e., the conversion of food into energy and growth - $\beta f_C h c_T w^q$) and ``active metabolism", the sum of standard (or resting) metabolism and the metabolic costs of foraging ($[1+\tau \delta] c_T k w^n$), with $\omega$ determining amount of oxygen required per unit of metabolised carbon. Metabolic scope is the difference between $S_{O_2}/w^n=f_{O_2}$, the amount of oxygen supplied per unit weight, and oxygen demand due to standard metabolism ($c_T k w^n$).

@@

We now assume that organisms will adjust their activity level to maximise available energy relative to mortality $M$: $\tau_{\text{opt}}$ = $\max_\tau(F)$ = $\max_\tau(P_C/M)$, with $F$ a measure of fitness \footnote{I \emph{think} this amounts to maximizing R0 if no more growth occurs: life expectancy is 1/M and reporoductive output is proportional to energy}. We further assume here that mortality scales with activity level and weight as $w^{q-1}$ (\cite{andersen_how_2009, hartvig_food_2011}), such that $M_{\tau,w} = \tau^\eta \rho w^{q-1}$, where $ \rho$ is mortality at mass $w=1$ and $\tau=1$, that is, at constant maximum activity and $M$ is the effective mortality at realised levels of activity at weight $w$.

@@

Life history theory suggests that the maximum\/asymptotic weight $W_{\infty}$ is found at the size at which all energy is allocated to reproduction, which optimally occurs at $\frac{\partial P_C/M}{\partial w}=1$ in the case of size-dependent mortality (provided concavity with respect to mass of this quantity). Note that $W_{\infty}$ depends on temperature only through $\tau$ and changes in $P_C$ as temperature changes metabolic and enzymatic rates. We thus impose no direct constraint on $W_{\infty}$.

@@

\begin{figure}[ht!]
\includegraphics[width=\linewidth]{activity} 
\caption[Activity, feeding and maximum length as a function of temperature]{Consequences of increasing temperature on activity levels, feeding and maximum weight for an individual weighing 100g. (a) Optimal (dashed line), limiting (dotted line) and realised (solid line) activity as a function of temperature. Realised activity is the smaller of the limiting and optimal levels at each temperature, with activity limited by O2 supply as pictured by dashed lines in b). Optimal acitivity steadily increases to compensate for rising metabolic costs (dotted line in b) with temperature until limited by oxygen (overlapping supply (dashed) and demand (solid) curves in b). (c) Rising activity lelves keeps feeding levels constant (solid line in c).  Mortality levels (dashed) increase as activity increases with temperature, while available energy (carbon, dotted line) for growth and reproduction peaks and decreases. Note that energy is only limited by oxygen-limited activity levels at low temperatures in scenario 3. The rate of change of available energy relative to mortality drives the adult size, which delines near linearly in log-space without oxygen limitations, and lower and upper limits imposed as oxygen limits activity and energy aquisition.}
\label{fig:act_fig}
\end{figure}

Higher temperature have cascading effects through their influence on both metabolic and ecological rates (Figure \ref{fig:act_fig}). To maintain optimal energy aquisition for growth and reproduction relative to mortality, activity levels generally increase, until they reach limitations imposed by oxygen supply (Figure \ref{fig:act_fig}a). This limit occurs when active metabolic costs equal oxygen supply (Figure \ref{fig:act_fig}b), and are therefore more pronounced for species with strongly dome-shaped maximum-metabolic rates (scenario 3 in Figure \ref{fig:act_fig}).

>>>>

Increasing temperatures under climate change are thought to affect individual physiology of ectotherms through increases in metabolic demands, leading to decreases in body-size with concomitant effects on species ecology. Although intuitively appealing, the magnitude, as well as when and where such changes are expected in an ecosystem context remains unclear. We develop a general trait-based model of carbon and oxygen budgets to determine the relative influence of thermal sensitivity, oxygen limitation, and environmental conditions on ectotherm body size.  Our model provides a general physiological basis for changes in species size and observed ecological consequences of temperature, and ties together various strands of thought about the ecological consequences of increased temperatures.

@@

Temperature, through its effects on individual physiology, is a dominant driver of species ecology and bio-geography (\cite{deutsch_climate_2015,pinsky_marine_2013,brown_toward_2004}). As a consequence, current and predicted temperature increases under climate change will act as a strong agent of change in many ecosystems (\cite{deutsch_climate_2015, stuart-smith_thermal_2015}). However, the nature of these changes can be difficult to predict as temperature effects scale from individuals to species and ecosystems. Through this cascade of scales, incorrect or approximate model assumptions at the individual scale can have disproportionate effects on ecosystem level outcomes (\cite{brander_overconfidence_2013, lefevre_models_2017}). In marine fish, recent predictions of decreasing organism size, and resulting decreases in the size of fisheries catches (\cite{cheung_shrinking_2013}), have been criticized as overly simplistic and not in line with physiological constraints (\cite{brander_overconfidence_2013, lefevre_models_2017}). While it is not disputed that organism size can change as a result of changing temperatures, the question remains when, where and why this is so.

Empirically, most ectotherms are found to grow larger in colder temperatures - a phenomenon called the temperature-size rule (\cite{atkinson_cells_2006}). Models of ectotherm growth can reproduce this pattern (under assumptions of optimal energy allocation between growth and reproduction) if increases in metabolic costs (in terms of energy) with temperature outpace increases in energy acquisition (\cite{kozlowski_can_2004}). However, while increasing metabolic costs are real, energy acquisition is an abstract, phenomenological description that lumps ecological aspects of foraging activity and prey availability. At a more mechanistic level of energy acquisition, consistent responses of foraging and energy acquisition with temperature have remained elusive (\cite{englund_temperature_2011,rall_universal_2012}), with variable responses across habitats and taxa. This is perhaps not surprising given variable foraging ecologies, but it remains unclear how temperature dependent feeding ecology impacts growth and adult size.

Physiologically, a long held view has been that temperature is a controlling factor while oxygen supply sets the physiological limits (\cite{fry_effects_1947,claireaux_linking_2007,lefevre_are_2016}). How exactly temperature influences physiological rates and limits, however, has been a matter of debate, not least because of the variable responses observed among different species. In most species, the standard metabolism (SM; the metabolic cost of maintenance and routine activity such as ventilation) increases near exponentially with temperature. A prevalent view is that the maximum metabolic rate (MMR; the metabolic rate at maximum sustained exercise) has a  dome-shaped response to temperature, whereby it can be increased (passively and actively) up to a point, but plateaus or decreases thereafter (\cite{fry_effects_1947,claireaux_linking_2007,lefevre_are_2016,portner_physiology_2008}). This leads to the view of a modal curve for metabolic scope (MMR-SM; the available oxygen/energy for additional activity), and suggests that towards the upper end of this curve, organisms will run out of oxygen.

This view was encapsulated in the Theory of Oxygen and Capacity Limitation of Temperature (\cite{portner_oxygen-and_2010}), which suggests that the decrease in metabolic scope towards extreme temperatures limits species ability to sustain core functions such as foraging and growth (i.e., functions beyond SM), and hence body size. In some species, however, maximum metabolic rate may be steadily increasing (\cite{lefevre_are_2016, verberk_does_2016}), suggesting that oxygen may not be the limiting factor at high temperatures. Consequently, the difference in observed physiological responses may at least partially determine temperature effects on body size for individual species. In addition to physiological differences, the assumed coefficients for temperature sensitivity of metabolism and size scaling exponents are different between species, and can disproportionately affect future organism size \cite{lefevre_models_2017, kozlowski_can_2004}.

@@

\section{Methods}

@@

Our model assumes that physiology is described by two key budgets: the energy and oxygen budgets. At a high level, we assume that animals will adapt activity levels to optimise available energy for growth and reproduction. Growth is limited either by food capture, food processing capacity, or by available oxygen. We further assume that temperature acts directly on rates that are determined by enzymatic activity: digestive activity (via maximum consumption) and metabolic costs. The scaling with temperature (determined by the activation energy $E_a$) is assumed the same for these processes. Consequently, temperature only acts on ecological rates (food acquisition) via optimisation given temperature driven changes in enzymatic rates.

@@

The oxygen budget $P_{0_2}$, determines physiological limits to aerobic activity. Demand ($D_{O_2}$) is the sum of oxygen used for SDA (i.e., the conversion of food into energy and growth - $\beta f_C h c_T w^q$) and ``active metabolism", the sum of standard (or resting) metabolism and the metabolic costs of foraging ($[1+\tau \delta] c_T k w^n$), with $\omega$ determining amount of oxygen required per unit of metabolised carbon. Metabolic scope is the difference between $S_{O_2}/w^n=f_{O_2}$, the amount of oxygen supplied per unit weight, and oxygen demand due to standard metabolism ($\omega c_T k w^n$).

@@

We now assume that organisms will adjust their activity level to maximise available energy relative to mortality $M$: $\tau_{\text{opt}}$ = $\max_\tau(F)$ = $\max_\tau(P_C/M)$, with $F$ a measure of fitness \footnote{I \emph{think} this amounts to maximizing R0 if no more growth occurs: life expectancy is 1/M and reproductive output is proportional to energy}. We further assume here that mortality scales with activity level and weight as $w^{q-1}$ (\cite{andersen_how_2009, hartvig_food_2011}), such that $M_{\tau,w} = \tau^\eta \rho w^{q-1}$, where $ \rho$ is mortality at mass $w=1$ and $\tau=1$, that is, at constant maximum activity and $M$ is the effective mortality at realised levels of activity at weight $w$.

@@

Life history theory suggests that the adult weight $W_{\infty}$ is found at the size at which all energy is allocated to reproduction, which optimally occurs at $\frac{\partial P_C/M}{\partial w}=1$ in the case of size-dependent mortality (provided concavity with respect to mass of this quantity). Note that $W_{\infty}$ depends on temperature only through $\tau$ and changes in $P_C$ as temperature changes metabolic and enzymatic rates. We thus impose no direct constraint on $W_{\infty}$.

\subsection{Trait based scenarios}

We now explore the sensitivity of size on a number of ecological and physiological traits. We model variable oxygen limitation through a set of three scenarios that reflect responses found empirically (\cite{lefevre_are_2016}; Figure \ref{fig:O2_fig}), ranging from an increasing MMR with temperature over most of the range (scenario 1), via a slightly dome-shaped response (scenario 2) to a strongly dome-shaped MMR (scenario 3).

Other physiological traits explored here are the FAS $\delta$, which determines the cost of active metabolism relative to standard metabolism. FAS ca vary strongly between species but is constant with weight (\cite{lefevre_models_2017, killen_ecological_2016}), and range from approximately 2 to 12, with most values found around 3-6. Here we explore values of 3-5. We explore the sensitivity to metabolic exponents by assuming that the exponent ranges between 0.85 and 0.9, with a base-case of 0.88 (\cite{lefevre_models_2017}).

@@

\begin{figure}[ht!]
\includegraphics[width=\linewidth]{activity} 
\caption[Activity, feeding and maximum length as a function of temperature]{Consequences of increasing temperature on activity levels, feeding and maximum weight for an individual weighing 100g. (a) Optimal (dashed line), limiting (dotted line) and realised (solid line) activity as a function of temperature. Realised activity is the smaller of the limiting and optimal levels at each temperature, with activity limited by O2 supply as pictured by dashed lines in b). Optimal activity steadily increases to compensate for rising metabolic costs (dotted line in b) with temperature until limited by oxygen (overlapping supply (dashed) and demand (solid) curves in b). (c) Rising activity levels keeps feeding levels constant (solid line in c).  Mortality levels (dashed) increase as activity increases with temperature, while available energy (carbon, dotted line) for growth and reproduction peaks and decreases. Note that energy is only limited by oxygen-limited activity levels at low temperatures in scenario 3. The rate of change of available energy relative to mortality drives the adult size, which declines near linearly in log-space without oxygen limitations, and lower and upper limits imposed as oxygen limits activity and energy acquisition.}
\label{fig:act_fig}
\end{figure}

Increasing temperatures have cascading effects through their influence on both metabolic and ecological rates (Figure \ref{fig:act_fig}). To maintain optimal energy acquisition for growth and reproduction relative to mortality, activity levels generally increase, until they reach limitations imposed by oxygen supply (Figure \ref{fig:act_fig}a). This limit occurs when active metabolic costs equal oxygen supply (Figure \ref{fig:act_fig}b), and are therefore more pronounced for species with strongly dome-shaped maximum-metabolic rates (scenario 3 in Figure \ref{fig:act_fig}).

### 2017-09-11 poster/poster.Rmd
final psoter etc

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = F,message = F,
                      cache=T,autodep = T,dev = "svg")  # Hide code by default. Can override chunk-by-chunk
```

@@

Size-at-maturation and maximum size are key traits, defining ecoloical niches and driving ecological interactions [@andersen_asymptotic_2006]. Temperature has been identified in various contexts as a key driver of organism size (Bergman's rule, Temperature size rule...) [@atkinson_temperature_1994]. Climate change may therefore have strong effects on organism size and follow on effects at population and ecosystem levels, but no clear framework to understand effects of temperature.

**Aim: Develop a eco-physiological model to predict effects of changing temperatures on organism size over ecological and evolutionary time-scales**

@@

@@

Feeding - type 2 response with parameters for maximum consumption rate $\gamma$ and maximum intake $h$, modulated by feeding activity $\tau$:
$$ f = \frac{{\color{uclagold}\tau} \gamma w^{p}\Theta}{{\color{uclagold}\tau}\gamma w^{p}\Theta  + {\color{bondiblue} T^c_{h} h w^{q}}} $$

Mass balance: Intake is assimilated with loss from heat increment $\phi$ and absorbtion $\beta$. Metabolic loss due to standard and feeding activity ($\tau$) metabolic costs
$$ P =(1-\beta-\phi)f {\color{bondiblue} T^c_{h} h w^{q}}-{\color{sinopia}k(1+{\color{uclagold}\tau\delta}) T^c_{M} w^n} $$

@@

## Ecological dynamics
With increasing temperature, most organisms will increase activity $\tau$ to offset increased metabolic costs.

Feeding level, available energy and mortality are impacted by response to temperature

Oxygen sets metabolic limits at extreme temperatures

# Sluggish small
ss <- list(beta=0.15,
           delta=4,
           Ea=0.52, 
           gamma=40, 
           h=20, 
           k=1.2,
           lO=1, 
           m=100, 
           M=0.1 ,
           n=0.88, 
           O2crit=2, 
           omega=0.25, 
           p=0.8,
           P50=4.2, 
           phi=0.25, 
           q=0.8 ,
           shape=3,
           temp=seq(5,26,l=100),
           Topt=20, 
           temp_ref=15,
           slope=-0.2,
           tmax=30,
           tr=0.3,
           v=6,
           lm=5000,
           nu=-0.2
)

# sluggish large
sl <- ss
sl$nu <- -0.3

# fast large
fl <- sl
fl$lO <- 2
fl$k <- 1.5
fl$h <- 40
fl$v <- 1
fl$M <- 0.3

# fast large
fs <- fl
fs$nu <- -0.2

scenarios <- bind_cols(data_frame(Scenario = as.factor(rep(c('Sluggish small','Sluggish large','Active small','Active large'),each=100))),
                       bind_rows(plot_data_temp(ss),
                                 plot_data_temp(sl),
                                 plot_data_temp(fs),
                                 plot_data_temp(fl)))

```{r dev.args= list(bg = 'transparent'),fig.width=7,fig.height=2}

ggplot(sc_act) + 
  geom_line(aes(x=Temperature, y=value, linetype=Activity),alpha=0.7) +
  theme_cowplot()+
  facet_wrap(~Scenario)+
  theme(strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank()) + 
  labs(x = expression("Mean temperature " ( degree*C)),
       y=expression('Activity ' (tau)))

```
\newline
```{r dev.args= list(bg = 'transparent'),fig.width=7,fig.height=2}

sc <- function(x) x/max(x,na.rm=T)

ggplot(sc_feed) + 
  geom_line(aes(x=Temperature, y=value, linetype=Rate),alpha=0.7) +
  theme_cowplot()+
  facet_wrap(~Scenario)+
  theme(strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank())+
  scale_linetype_discrete('')+
  labs(x = expression("Mean temperature " ( degree*C)),
       y='Relative rate')

Temperature affects size via evolved reaction norm on ecological time-scales:

```{r dev.args= list(bg = 'transparent'),fig.width=7,fig.height=2}

gout_fs <- plot_data_growth_tO2(fs,n_int = 50)

#browser()
norm <- gout_fs$alloc %>% 
  group_by(Temperature) %>%
  summarise(ts=t[ifelse(any(abs(allocs-0.5)<0.1),which.min(abs(allocs-0.5)),NA)-1],
            m=ls[ifelse(any(abs(allocs-0.5)<0.1),which.min(abs(allocs-0.5)),NA)-1])

cp <- ggplot() +
  geom_line(aes(x=t, y=ls,col=as.factor(Temperature)),data=gout_fs$alloc) +
  geom_point(aes(x=ts, y=m),data=norm) +
  #geom_point(data=growth_scenarios[seq.int(1,nrow(growth_scenarios),by = 5),],aes(x=Temperature, y=Temp, col=Scenario,shape=Scenario),alpha=0.5,size=2) +
  #theme_cowplot()+
  viridis::scale_colour_viridis(discrete = T,guide='none')+
  theme(
    panel.background = element_blank(),
    plot.background = element_blank())+
  coord_cartesian()+
  labs(y = 'Length (cm)',
       x='Age  (years)')

scs <- list(ss,sl,fs,fl)
scn <- unique(scenarios$Scenario)
gouts <- lapply(scs,plot_data_growth_tO2,n_int = 50)

alloc <- bind_rows(lapply(1:length(gouts), function(x) {
  y <- gouts[[x]][['alloc']]
  y$Scenario <- scn[x]
  y}))

norm <- alloc %>% 
  group_by(Scenario,Temperature) %>%
  summarise(ts=t[ifelse(any(abs(allocs-0.5)<0.05),which.min(abs(allocs-0.5)),NA)-1],
            ls=ls[ifelse(any(abs(allocs-0.5)<0.05),which.min(abs(allocs-0.5)),NA)-1],
            alloc=allocs[ifelse(any(abs(allocs-0.5)<0.05),which.min(abs(allocs-0.5)),NA)-1]) %>%
  mutate(ls = sc(ls))

np <-  ggplot() +
  geom_line(aes(x=Temperature, y=ls, linetype=Scenario),data=norm) +
  theme(
    panel.background = element_blank(),
    plot.background = element_blank())+
  #scale_color_discrete(guide='none')+
  #scale_shape_discrete(guide='none')+
  coord_cartesian()+
  labs(y = 'Relative length',
       x='Temperature')

@@

## Traits

Two strategists: minimising mortality $M$ or maximising energy gain $P$

$P$ strategy: 
* Maximise intake at the cost of higher standard metabolism to feeding maintain activity and larger gut.
* Live with higher $M$

$M$ strategy: 
* Minimise $M$ by minimising dangerous feeding activty, minimise resting metabolism

## Evolutionary impact

Size-at maturation reacts strongly to temperature changes in the model, however, changes are small on ecological time-scales ($\lt$1mm per year)

```{r dev.args= list(bg = 'transparent'),fig.width=7,fig.height=3}

@@

wtfs <- winfs %>% group_by(Scenario) %>%
  mutate(length=sc(lw(Temp)))

cpG <- ggplot(wtfs) +
  geom_line(aes(x=Temperature, y=length, linetype=Scenario),alpha=0.7) +
  #geom_point(data=growth_scenarios[seq.int(1,nrow(growth_scenarios),by = 5),],aes(x=Temperature, y=Temp, col=Scenario,shape=Scenario),alpha=0.5,size=2) +
  #theme_cowplot()+
  scale_linetype_discrete(guide='none')+
  theme(
    panel.background = element_blank(),
    plot.background = element_blank())+
  #scale_color_discrete(guide='none')+
  #scale_shape_discrete(guide='none')+
  #scale_y_log10()+
  labs(x = expression("Mean temperature " ( degree*C)),
       y='Relative adult length')

npG <- ggplot(wtfs) +
  geom_line(aes(x=Temperature, y=L/length, linetype=Scenario),alpha=0.7) +
  #geom_point(data=growth_scenarios[seq.int(1,nrow(growth_scenarios),by = 5),],aes(x=Temperature, y=Temp, col=Scenario,shape=Scenario),alpha=0.5,size=2) +
  #theme_cowplot()+
  theme(
    panel.background = element_blank(),
    plot.background = element_blank())+
  #scale_color_discrete(guide='none')+
  #scale_shape_discrete(guide='none')+
  coord_cartesian(ylim=quantile(wtfs$L/wtfs$length,c(0.05,0.95),na.rm=T)) + 
  labs(x = expression("Mean temperature " ( degree*C)),
       y=expression(Percent~change~y^-1))

@@

@@

## Acknowledgements

P. Neubauer acknowledges funding from the Royal Society of New Zealand Marsden Fast Start grant 14-DFG-001.

All code for this poster is accessble from github.

# {.col-1}
## References

>>>>

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,warning = F,message = F,
                      cache=T,autodep = T,dev = "svg",dev.args= list(bg  = 'transparent'))  # Hide code by default. Can override chunk-by-chunk
```

int_m <- 1000
int_temp <- 40

# Sluggish small
ss <- list(c=0.3,
           beta=0.15,
           delta=4,
           Ea=0.52, 
           gamma=40, 
           h=20, 
           k=1.2,
           lO=1, 
           m=100, 
           M=0.1 ,
           n=0.88, 
           O2crit=2, 
           omega=0.25, 
           p=0.8,
           P50=4.2, 
           phi=0.25, 
           q=0.8 ,
           shape=3,
           temp=seq(5,26,l=int_temp),
           Topt=20, 
           temp_ref=15,
           slope=-0.2,
           tmax=30,
           tr=0.9,
           v=6,
           lm=int_m,
           nu=-0.2
)

# fast small
fs <- ss
fs$lO <- 2
fs$k <- 1.5
fs$h <- 40
fs$v <- 6
fs$M <- 0.3
fs$tmax <- 20
fs$nu <- -0.2

scenarios <- bind_cols(data_frame(Scenario = as.factor(rep(c('M strategy','P strategy'),each=int_temp))),
                       bind_rows(plot_data_temp(ss),
                                 plot_data_temp(fs)))

@@

&nbsp;

Size-at-maturation and maximum size are key traits, defining ecological niches and driving ecological interactions [@andersen_asymptotic_2006]. Temperature has been identified in various contexts as a key driver of organism size (Bergman's rule, Temperature-size-rule...) [@atkinson_temperature_1994]. Climate change may therefore have strong effects on organism size and follow on effects at population and ecosystem levels, but no clear framework to understand effects of temperature [@lefevre_models_2017].

&nbsp;

**Aim: Develop a eco-physiological model [@holt_climate_2014] to predict effects of changing temperatures on organism size over ecological and evolutionary time-scales**

@@

&nbsp;

@@

Feeding - type 2 functional response with parameters for maximum consumption rate $\gamma$, modulated by feeding activity $\tau$, and maximum intake $h$, which responds to temperature via $T^c_{M}$:
$$ f = \frac{{\color{uclagold}\tau} \gamma w^{p}\Theta}{{\color{uclagold}\tau}\gamma w^{p}\Theta  + {\color{bondiblue} T^c_{h} h w^{q}}} $$

Mass balance: Intake is assimilated with loss from heat increment $\phi$ and absorption $\beta$. Metabolic loss due to standard and feeding activity ($\tau$) metabolic costs, modulated by temperature via $T^c_{M}$
$$ P =(1-\beta-\phi)f {\color{bondiblue} T^c_{h} h w^{q}}-{\color{sinopia}k(1+{\color{uclagold}\tau\delta}) T^c_{M} w^n} $$

@@

&nbsp;

Strategy trade-off: minimising mortality $M$ or maximising energy gain $P$

&nbsp;

$P$ strategy: Maximise intake at the cost of higher standard metabolism to maintain feeding activity and larger gut. Live with higher $M$.

&nbsp;

$M$ strategy: Minimise $M$ by minimising dangerous feeding activity, minimise resting metabolism.

## Temperature impact on size

&nbsp;

```{r fig.width=7,fig.height=2.5,fig.cap='Figure 3: **On ecological time-scales**, temperature affects size via evolved maturation-reaction norms [@marty_impact_2011]. Most reaction norms found in nature have a negative slope, meaning for faster growth, maturation occurs at a larger size. Increasing temparture (left panel, purple to yellow growth curves) modifies growth, and maturation age changes according to the reaction norm (right panel).'}

gout_fs <- plot_data_growth_tO2(fs)

#browser()
norm <- gout_fs$t_growth %>% 
  group_by(Temperature) %>%
  summarise(ts=t[ifelse(any(abs(allocs-0.5)<0.1),which.min(abs(allocs-0.5)),NA)-1],
            m=t_length[ifelse(any(abs(allocs-0.5)<0.1),which.min(abs(allocs-0.5)),NA)-1])

cp <- ggplot() +
  geom_line(aes(x=t, y=t_length,col=as.factor(Temperature)),data=gout_fs$t_growth) +
  geom_point(aes(x=ts, y=m),data=norm) +
  #geom_point(data=growth_scenarios[seq.int(1,nrow(growth_scenarios),by = 5),],aes(x=Temperature, y=Temp, col=Scenario,shape=Scenario),alpha=0.5,size=2) +
  theme_cowplot()+
  viridis::scale_colour_viridis(discrete = T,guide='none')+
  theme(plot.margin = margin(2, 6, 0, 6),
        panel.background = element_blank(),
        plot.background = element_blank())+
  coord_cartesian()+
  labs(y = 'Length (cm)',
       x='Age  (years)')

scs <- list(ss,fs)
scn <- unique(scenarios$Scenario)
gouts <- parallel::mclapply(scs,plot_data_growth_tO2,mc.cores = 4)

alloc <- bind_rows(lapply(1:length(gouts), function(x) {
  y <- gouts[[x]][['t_growth']]
  y$Scenario <- scn[x]
  y}))

norm <- alloc %>% 
  group_by(Scenario,Temperature) %>%
  summarise(ts=t[ifelse(any(abs(allocs-0.5)<0.05),which.min(abs(allocs-0.5)),NA)-1],
            ls=t_length[ifelse(any(abs(allocs-0.5)<0.05),which.min(abs(allocs-0.5)),NA)-1],
            alloc=allocs[ifelse(any(abs(allocs-0.5)<0.05),which.min(abs(allocs-0.5)),NA)-1]) %>%
  mutate(ls = sc(ls))

np <-  ggplot() +
  geom_line(aes(x=Temperature, y=ls, linetype=Scenario),data=norm) +
  theme_cowplot()+
  theme(plot.margin = margin(2, 6, 0, 6),
        panel.background = element_blank(),
        plot.background = element_blank())+
  #scale_color_discrete(guide='none')+
  #scale_shape_discrete(guide='none')+
  coord_cartesian()+
  labs(y = 'Relative length',
       x='Temperature')

@@

&nbsp;

```{r fig.width=7,fig.height=2.5,fig.cap='Figure 4: **On evolutionary time-scales**, selection moves the reaction norm. Size-at-maturation reacts strongly to temperature changes in the model (left panel). Selection (S) induced changes in size are slow for M strategists, fast for P strategists.'}

@@

wtfs <- winfs %>% group_by(Scenario) %>%
  mutate(length=sc(lw(l)))

cpG <- ggplot(wtfs) +
  geom_line(aes(x=Temperature, y=length, linetype=Scenario),alpha=0.7) +
  theme_cowplot()+
  scale_linetype_discrete(guide='none')+
  theme(plot.margin = margin(3, 6, 0, 6),
        panel.background = element_blank(),
        plot.background = element_blank())+
  labs(x = expression("Mean temperature " ( degree*C)),
       y='Relative adult length')

npG <- ggplot(wtfs) +
  geom_line(aes(x=Temperature, y=L/length, linetype=Scenario),alpha=0.7) +
  theme(plot.margin = margin(3, 6, 0, 6),
        panel.background = element_blank(),
        plot.background = element_blank())+
  coord_cartesian(ylim=quantile(wtfs$L/wtfs$length,c(0.05,0.95),na.rm=T)) + 
  labs(x = expression("Mean temperature " ( degree*C)),
       y=expression(S~response~y^-1))

@@

## Ecological dynamics

```{r fig.width=7,fig.height=2,fig.cap='Figure 1: With increasing temperature, most organisms will increase activity $\\tau$ to offset increased metabolic costs. Oxygen sets metabolic limits at extreme temperatures.'}

ggplot(sc_act) + 
  geom_line(aes(x=Temperature, y=value, linetype=Activity),alpha=0.7) +
  theme_cowplot()+
  facet_wrap(~Scenario)+
  theme(plot.margin = margin(0, 6, 0, 6),
        strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank()) + 
  labs(x = expression("Mean temperature " ( degree*C)),
       y=expression('Activity ' (tau)))

```{r fig.width=7,fig.height=2,fig.cap='Figure 2: Feeding level, available energy and mortality are impacted by increased activity and the metabolic response to temperature.'}

ggplot(sc_feed) + 
  geom_line(aes(x=Temperature, y=value, linetype=Rate),alpha=0.7) +
  theme_cowplot()+
  facet_wrap(~Scenario)+
  theme(plot.margin = margin(0, 6, 0, 6),
        strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank())+
  scale_linetype_discrete('')+
  labs(x = expression("Mean temperature " ( degree*C)),
       y='Relative rate')

## Take home messages

<span style="font-size: 300%;">

&nbsp;

&nbsp;

* **On ecological time-scales changes in adult size due to temperature alone will likely be small, and masked by changes in the environment.**

&nbsp;

&nbsp;

* **On evolutionary time-scales, size responds strongly. The relative selection response is strongest at temperature extremes.**

&nbsp;

&nbsp;

* **Size of $P$ strategy species with high activity reacts more strongly compared to sluggish $M$ strategy species.**

</span>

@@

## Acknowledgements & References

P. Neubauer acknowledges funding from the Royal Society of New Zealand Marsden Fast Start grant 14-DFG-001. All code for the model and this poster is accessible from github at repository Adaptive-dynamics/Sizing-fx-of-temperature.

&nbsp;
